Add Plant Functional Type (PFT) labels

### 2021-04-22 R/add_plant_functional_type.Rmd
working through taxons

- common name
- pft_MRA (based on Macrosystems proposed labels)
    - coniferous_forest
    - deciduous_forest
    - woody_shrub
- pft_MODIS (based on MODIS land cover classes)
    - evergreen_needleleaf
    - evergreen_broadleaf
    - deciduous_needlelead
    - deciduous_broadleaf
    - shrub
- phenology (evergreen or deciduous)
- leaf type (broadleaf or needleleaf)
- growth form (shrub or tree, based on NEON field based observations)

```{r manually_add_pft_labels}
# list unique taxon labels, add an empty columns to fill with plant functional type (PFT) and common name labels.
taxon_pft_lookup <- data.frame(genusSpecies = unique(cleaned_spectra$genusSpecies),
                               commonName = NA,
                               pft_MRA = NA,
                               pft_MODIS = ,
                               phenology = NA,
                               leafType = NA,
                               growthForm = NA) %>% 
  # arrange alphabetically
  dplyr::arrange(genusSpecies)

@@

  if(genusSpecies == "Abies amabilis"){
    taxon_pft_lookup$pft_MRA[i] = "coniferous_forest" 
    taxon_pft_lookup$pft_MODIS[i] = 
    taxon_pft_lookup$phenology[i] = "evergreen"
    taxon_pft_lookup$leafType[i] = "needleleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "Pacific silver fir"}

  else if(genusSpecies == "Abies balsamea"){
    taxon_pft_lookup$pft_MRA[i] = "coniferous_forest" 
    taxon_pft_lookup$phenology[i] = "evergreen"
    taxon_pft_lookup$leafType[i] = "needleleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "balsam fir"}

  else if(genusSpecies == "Abies lasiocarpa"){
    taxon_pft_lookup$pft_MRA[i] = "coniferous_forest"
    taxon_pft_lookup$phenology[i] = "evergreen"
    taxon_pft_lookup$leafType[i] = "needleleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "subalpine fir"}

  else if(genusSpecies == "Acacia greggii"){
    taxon_pft_lookup$pft_MRA[i] = "woody_shrub" 
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = ""
    taxon_pft_lookup$growthForm[i] = "shrub"
    taxon_pft_lookup$commonName[i] = "catclaw acacia"}

  else if(genusSpecies == "Acer barbatum"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "Southern sugar maple"}

  else if(genusSpecies == "Acer circinatum"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest" 
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "vine maple"}

  else if(genusSpecies == "Acer negundo"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "boxelder"}

  else if(genusSpecies == "Acer pensylvanicum"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "striped maple"}

  else if(genusSpecies == "Acer rubrum"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "red maple"}

  else if(genusSpecies == "Acer saccharinum"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "silver maple"}

  else if(genusSpecies == "Acer saccharum"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "sugar maple"}

  else if(genusSpecies == "Ailanthus altissima"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "tree of heaven"}

  else if(genusSpecies == "Alnus rubra"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "red alder"}

  else if(genusSpecies == "Aloysia wrightii"){
    taxon_pft_lookup$pft_MRA[i] = "woody_shrub"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "NA"
    taxon_pft_lookup$growthForm[i] = "shrub"
    taxon_pft_lookup$commonName[i] = "Wright's beebrush"}

  else if(genusSpecies == "Amelanchier laevis"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "Allegheny serviceberry"}

  else if(genusSpecies == "Amyris elemifera"){
    taxon_pft_lookup$pft_MRA[i] = "???" # is not deciduous or coniferous forest
    taxon_pft_lookup$phenology[i] = "evergreen"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "sea torchwood"}

  else if(genusSpecies == "Artemisia tridentata"){
    taxon_pft_lookup$pft_MRA[i] = "woody_shrub"
    taxon_pft_lookup$phenology[i] = "evergreen"
    taxon_pft_lookup$leafType[i] = "NA"
    taxon_pft_lookup$growthForm[i] = "shrub"
    taxon_pft_lookup$commonName[i] = "big sagebrush"}

  else if(genusSpecies == "Atriplex canescens"){
    taxon_pft_lookup$pft_MRA[i] = "woody_shrub"
    taxon_pft_lookup$phenology[i] = "evergreen"
    taxon_pft_lookup$leafType[i] = "NA"
    taxon_pft_lookup$growthForm[i] = "shrub"
    taxon_pft_lookup$commonName[i] = "fourwing saltbush"}

  else if(genusSpecies == "Atriplex confertifolia"){
    taxon_pft_lookup$pft_MRA[i] = "woody_shrub"
    taxon_pft_lookup$phenology[i] = "evergreen"
    taxon_pft_lookup$leafType[i] = "NA"
    taxon_pft_lookup$growthForm[i] = "shrub"
    taxon_pft_lookup$commonName[i] = "shadscale saltbush"}

  else if(genusSpecies == "Betula alleghaniensis"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "yellow birch"}

  else if(genusSpecies == "Betula lenta"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "sweet birch"}

  else if(genusSpecies == "Betula neoalaskana"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "resin birch"}

  else if(genusSpecies == "Betula nigra"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "river birch"}

  else if(genusSpecies == "Betula papyrifera"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "paper birch"}

  else if(genusSpecies == "Betula sp."){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "birch"}

  else if(genusSpecies == "Bourreria sp."){
    taxon_pft_lookup$pft_MRA[i] = "NA"  # taxon class is too broad to specify 
    taxon_pft_lookup$phenology[i] = "NA"
    taxon_pft_lookup$leafType[i] = "NA"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "strongbark"}

  else if(genusSpecies == "Bursera simaruba"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"  
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "gumbo limbo"}

  else if(genusSpecies == "Carpinus caroliniana"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "American hornbeam"}

  else if(genusSpecies == "Carya aquatica"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "water hickory"}

  else if(genusSpecies == "Carya cordiformis"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "bitternut hickory"}

  else if(genusSpecies == "Carya glabra"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "pignut hickory"}

  else if(genusSpecies == "Carya illinoinensis"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "pecan"}

  else if(genusSpecies == "Carya ovalis"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "red hickory"}

  else if(genusSpecies == "Carya ovata"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "shagbark hickory"}

  else if(genusSpecies == "Carya tomentosa"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "mockernut hickory"}

  else if(genusSpecies == "Celtis laevigata"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "sugarberry"}

  else if(genusSpecies == "Celtis occidentalis"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "common hackberry"}

  else if(genusSpecies == "Cercis canadensis"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "eastern redbud"}

  else if(genusSpecies == "Chrysothamnus viscidiflorus"){
    taxon_pft_lookup$pft_MRA[i] = "woody_shrub"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "NA"
    taxon_pft_lookup$growthForm[i] = "shrub"
    taxon_pft_lookup$commonName[i] = "yellow rabbitbrush"}

  else if(genusSpecies == "Cordia collococca"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "red manjack"}

  else if(genusSpecies == "Cornus drummondii"){
    taxon_pft_lookup$pft_MRA[i] = "woody_shrub" 
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "NA"
    taxon_pft_lookup$growthForm[i] = "shrub"
    taxon_pft_lookup$commonName[i] = "roughleaf dogwood"}

  else if(genusSpecies == "Cornus florida"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "flowering dogwood"}

  else if(genusSpecies == "Corylus cornuta"){
    taxon_pft_lookup$pft_MRA[i] = "woody_shrub" 
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = ""
    taxon_pft_lookup$growthForm[i] = "shrub"
    taxon_pft_lookup$commonName[i] = "beaked hazelnut"}

  else if(genusSpecies == "Diospyros virginiana"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "common persimmon"}

>>>>

- common name
- pft_MRA (based on Macrosystems proposed labels)
    - coniferous_forest
    - deciduous_forest
    - woody_shrub
- pft_MODIS (based on MODIS land cover classes)
    - evergreen-needleleaf_forest
    - evergreen-broadleaf_forest
    - deciduous-needleleaf_forest
    - deciduous-broadleaf_forest
    - shrub
- phenology (evergreen or deciduous)
- leaf type (broadleaf or needleleaf)
- growth form (shrub or tree, based on NEON field based observations)

```{r manually_add_pft_labels}
# list unique taxon labels, add an empty columns to fill with plant functional type (PFT) and common name labels.
taxon_pft_lookup <- data.frame(genusSpecies = unique(cleaned_spectra$genusSpecies),
                               commonName = NA,
                               pft_MRA = NA,
                               pft_MODIS = NA,
                               phenology = NA,
                               leafType = NA,
                               growthForm = NA) %>% 
  # arrange alphabetically
  dplyr::arrange(genusSpecies)

@@

  if(genusSpecies == "Abies amabilis"){
    taxon_pft_lookup$pft_MRA[i] = "coniferous_forest" 
    taxon_pft_lookup$pft_MODIS[i] = "evergreen-needleleaf_forest"
    taxon_pft_lookup$phenology[i] = "evergreen"
    taxon_pft_lookup$leafType[i] = "needleleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "Pacific silver fir"}

  else if(genusSpecies == "Abies balsamea"){
    taxon_pft_lookup$pft_MRA[i] = "coniferous_forest" 
    taxon_pft_lookup$pft_MODIS[i] = "evergreen-needleleaf_forest"
    taxon_pft_lookup$phenology[i] = "evergreen"
    taxon_pft_lookup$leafType[i] = "needleleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "balsam fir"}

  else if(genusSpecies == "Abies lasiocarpa"){
    taxon_pft_lookup$pft_MRA[i] = "coniferous_forest"
    taxon_pft_lookup$pft_MODIS[i] = "evergreen-needleleaf_forest"
    taxon_pft_lookup$phenology[i] = "evergreen"
    taxon_pft_lookup$leafType[i] = "needleleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "subalpine fir"}

  else if(genusSpecies == "Acacia greggii"){
    taxon_pft_lookup$pft_MRA[i] = "woody_shrub" 
    taxon_pft_lookup$pft_MODIS[i] = "shrub"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "NA"
    taxon_pft_lookup$growthForm[i] = "shrub"
    taxon_pft_lookup$commonName[i] = "catclaw acacia"}

  else if(genusSpecies == "Acer barbatum"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$pft_MODIS[i] = "deciduous-broadleaf_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "Southern sugar maple"}

  else if(genusSpecies == "Acer circinatum"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest" 
    taxon_pft_lookup$pft_MODIS[i] = "deciduous-broadleaf_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "vine maple"}

  else if(genusSpecies == "Acer negundo"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$pft_MODIS[i] = "deciduous-broadleaf_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "boxelder"}

  else if(genusSpecies == "Acer pensylvanicum"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$pft_MODIS[i] = "deciduous-broadleaf_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "striped maple"}

  else if(genusSpecies == "Acer rubrum"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$pft_MODIS[i] = "deciduous-broadleaf_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "red maple"}

  else if(genusSpecies == "Acer saccharinum"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$pft_MODIS[i] = "deciduous-broadleaf_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "silver maple"}

  else if(genusSpecies == "Acer saccharum"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$pft_MODIS[i] = "deciduous-broadleaf_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "sugar maple"}

  else if(genusSpecies == "Ailanthus altissima"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$pft_MODIS[i] = "deciduous-broadleaf_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "tree of heaven"}

  else if(genusSpecies == "Alnus rubra"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$pft_MODIS[i] = "deciduous-broadleaf_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "red alder"}

  else if(genusSpecies == "Aloysia wrightii"){
    taxon_pft_lookup$pft_MRA[i] = "woody_shrub"
    taxon_pft_lookup$pft_MODIS[i] = "shrub"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "NA"
    taxon_pft_lookup$growthForm[i] = "shrub"
    taxon_pft_lookup$commonName[i] = "Wright's beebrush"}

  else if(genusSpecies == "Amelanchier laevis"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$pft_MODIS[i] = "deciduous-broadleaf_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "Allegheny serviceberry"}

  else if(genusSpecies == "Amyris elemifera"){
    taxon_pft_lookup$pft_MRA[i] = "NA" # is not deciduous or coniferous forest
    taxon_pft_lookup$pft_MODIS[i] = "evergreen-broadleaf_forest"
    taxon_pft_lookup$phenology[i] = "evergreen"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "sea torchwood"}

  else if(genusSpecies == "Artemisia tridentata"){
    taxon_pft_lookup$pft_MRA[i] = "woody_shrub"
    taxon_pft_lookup$pft_MODIS[i] = "shrub"
    taxon_pft_lookup$phenology[i] = "evergreen"
    taxon_pft_lookup$leafType[i] = "NA"
    taxon_pft_lookup$growthForm[i] = "shrub"
    taxon_pft_lookup$commonName[i] = "big sagebrush"}

  else if(genusSpecies == "Atriplex canescens"){
    taxon_pft_lookup$pft_MRA[i] = "woody_shrub"
    taxon_pft_lookup$pft_MODIS[i] = "shrub"
    taxon_pft_lookup$phenology[i] = "evergreen"
    taxon_pft_lookup$leafType[i] = "NA"
    taxon_pft_lookup$growthForm[i] = "shrub"
    taxon_pft_lookup$commonName[i] = "fourwing saltbush"}

  else if(genusSpecies == "Atriplex confertifolia"){
    taxon_pft_lookup$pft_MRA[i] = "woody_shrub"
    taxon_pft_lookup$pft_MODIS[i] = "shrub"
    taxon_pft_lookup$phenology[i] = "evergreen"
    taxon_pft_lookup$leafType[i] = "NA"
    taxon_pft_lookup$growthForm[i] = "shrub"
    taxon_pft_lookup$commonName[i] = "shadscale saltbush"}

  else if(genusSpecies == "Betula alleghaniensis"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$pft_MODIS[i] = "deciduous-broadleaf_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "yellow birch"}

  else if(genusSpecies == "Betula lenta"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$pft_MODIS[i] = "deciduous-broadleaf_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "sweet birch"}

  else if(genusSpecies == "Betula neoalaskana"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$pft_MODIS[i] = "deciduous-broadleaf_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "resin birch"}

  else if(genusSpecies == "Betula nigra"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$pft_MODIS[i] = "deciduous-broadleaf_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "river birch"}

  else if(genusSpecies == "Betula papyrifera"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$pft_MODIS[i] = "deciduous-broadleaf_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "paper birch"}

  else if(genusSpecies == "Betula sp."){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$pft_MODIS[i] = "deciduous-broadleaf_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "birch"}

  else if(genusSpecies == "Bourreria sp."){
    taxon_pft_lookup$pft_MRA[i] = "NA"  # taxon class is too broad to specify 
    taxon_pft_lookup$pft_MODIS[i] = "NA"
    taxon_pft_lookup$phenology[i] = "NA"
    taxon_pft_lookup$leafType[i] = "NA"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "strongbark"}

  else if(genusSpecies == "Bursera simaruba"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"  
    taxon_pft_lookup$pft_MODIS[i] = "deciduous-broadleaf_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "gumbo limbo"}

  else if(genusSpecies == "Carpinus caroliniana"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$pft_MODIS[i] = "deciduous-broadleaf_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "American hornbeam"}

  else if(genusSpecies == "Carya aquatica"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$pft_MODIS[i] = "deciduous-broadleaf_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "water hickory"}

  else if(genusSpecies == "Carya cordiformis"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$pft_MODIS[i] = "deciduous-broadleaf_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "bitternut hickory"}

  else if(genusSpecies == "Carya glabra"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$pft_MODIS[i] = "deciduous-broadleaf_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "pignut hickory"}

  else if(genusSpecies == "Carya illinoinensis"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$pft_MODIS[i] = "deciduous-broadleaf_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "pecan"}

  else if(genusSpecies == "Carya ovalis"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$pft_MODIS[i] = "deciduous-broadleaf_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "red hickory"}

  else if(genusSpecies == "Carya ovata"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$pft_MODIS[i] = "deciduous-broadleaf_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "shagbark hickory"}

  else if(genusSpecies == "Carya tomentosa"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$pft_MODIS[i] = "deciduous-broadleaf_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "mockernut hickory"}
